Clicks and Cliques

### 2018-12-23 papersoul.Rnw
minor edit

# to be removed...
library(reshape2)
library(plyr)

@@

<<data,echo=FALSE,warning=FALSE,message=FALSE, cache=FALSE>>=
# Procesing data sets from SOCT and Census Bureau
source('code/data_chk_tidy.R')
source('code/census_chk_tidy.R')
# loading results from cluster and random forest models.

res_cluster <- readRDS('data/res_cluster.rds') %>%
list2env(globalenv())

@@

# Compute comunity proportion of attached people
# community labels differ from soul and census datasets

lev <- soul %>% pull(QSB) %>% levels()

@@

rankings <- important %>%
  filter(rnk < 6) %>%
  group_by(dim, rnk) %>% 
  summarize(V1 = length(dim)) %>%
  ungroup() %>% 
  mutate(rnk = factor(rnk, labels = paste("R", 1:5, sep = "")))

@

@@

ggplot() + 
  geom_path(data = us, aes(long, lat, group = group), color = 'grey' ) +
  geom_point(data = lon.lat, aes(x = lon, y = lat, size = pop100), alpha = .5) +
  xlab('') + ylab('') + 
  scale_size(range = c(5,20)) + 
  scale_color_brewer(type = 'seq') +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(), 
        legend.position = 'none')
@
\caption{Geographical location of the 26 communities in the data, dot's size represents population. Only two communities on the west coast are included. \label{map}}
\end{figure}

@@

\begin{figure}[hbpt]
<<corplot, echo=FALSE, dependson=c('data'),cache=TRUE, warning = FALSE>>=
# With Census data we want to describe communities

cors <- census.data %>% 
  gather(vars, vv, -QSB, -CCEGRP2Attached, -`CCEGRP2Not Attached` ) %>%
  group_by(vars) %>% 
  summarise(
    cor.att = cor(vv, CCEGRP2Attached, use = 'complete.obs'),
    cor.noatt = cor(vv, `CCEGRP2Not Attached`, use = 'complete.obs') ) %>% 
  filter(!is.na(cor.att), vars != 'unemployed') %>% 
  filter(rank(cor.att) %in% c(1:10, 52:61) ) %>% 
  mutate(vars =  fct_reorder(vars, cor.att) , vars = recode_factor(
      vars, 
      "av.own" = 'Aver.years.own.in.community',
      "pop100"= 'size.community', 
      "av.rent"= 'Average.years.rent.in.community',
      "ow.prop10" = 'Prop.owner.more.10.years',
      "prop.female" = 'Pro.female',
      "prop2.1worker" = 'Prop.house.2.per.1.worker',
      "prop.4.per" = 'Prop.house.4.per.or.more',
      "ren.prop10" = 'Prop.rent.more.10.years',
      "own.rate" = 'Prop.own',
      "un.rate" = 'Unemployment.rate',
      "prop1.worker" = 'Prop.1worker',
      "Bachelor.s.degree.or.higher." = 'Prop.bachelor',
      "prop3.3worker" = 'Prop.house.3.per.3.workers',
      "prop.nat.entry.90s" = 'Prop.native.entry.90s',
      "prop.nat.entry.80s" = 'Prop.native.entry.80s',
      "militar" = 'Prop.militar',
      "prop1821" = 'Prop.pop.bw.18.21',
      "prop.entry.80s" = 'Prop.new.entry',
      "prop.male" = 'Prop.male',
      "prop.2.per" = 'Prop.2perhouse')) %>% 
  gather(type, vv, -vars) %>% 
  mutate( type = factor(type, labels = c('Attached', 'Not Attached') ),
          positive = factor(sign(vv), labels = c( 'Negative', 'Positive') ) )

cors %>%
  ggplot() +
  geom_point(aes(y = vars, x = vv, color = positive), size = I(3)) +
  facet_grid(~type) + theme(legend.position = 'bottom', legend.title = element_blank(), axis.text.y = element_text(size = I(15))) + labs(y = "", x = "Correlation")

@
\caption{Correlations with attached and not attached proportion in order from top to bottom based on correlation with attached\label{corplot}. This plot shows the 10 more relevant variables with positive correlation and 10 with negative correlation with attachment.}
\end{figure}

@@

cors %>%  
  filter(vars %in% c("Prop.new.entry", "Average.years.rent.in.community", "Aver.years.own.in.community") &  type =='Attached' ) %>% 
  select(vars, vv) %>% rename(cor = vv) %>%
  bind_rows(data.frame(vars = 'stay', cor = rho)) %>% 
  mutate(Data =  c(rep('Census Bureau', 3), 'Knight-Gallup data'), 
         Description = c( 'Average years owning the current home', 
                          'Average years renting in current home', 
                          'Prop. of people not borned in the community', 
                          'Proportion of age in current comunity')) %>% 
  rename(Correlation = cor) %>% 
  select(Data, Description, Correlation) %>% 
  xtable( caption='Correlation with attachment', label='lifetab') %>% 
  print( caption.placement = 'top', include.rownames = FALSE)

@

@@

\begin{figure}
<<life_seq, dependson=c('life'), echo=FALSE,fig.height=4,message=FALSE,warning=FALSE>>=
life %<>% life %>%
  mutate(QSB = factor(QSB, levels = rev(levels(QSB))))

@@

aux <- soul$QD111

levels(aux)[c(4,6:10,15:17,19,20) ] <- 'Other'
levels(aux)[c(1,2,5,7,8,10) ] <- 'Non Response'
levels(aux)[2] <- 'Hispanic'
aux[soul$QD10 == 'Yes'] <- 'Hispanic'
soul$race  <- factor(aux)

@@

ggplot(data = tab, aes(x = diver, y = logit(statistic.CCEGRP2Attached))) + geom_point(size = I(5)) + theme(axis.text.y = element_text( size = I(10)) ) + labs( y = 'Proportion of Attach (logit scale)', x = 'Race diversity index') + geom_smooth(se = FALSE, method = 'lm', size = I(1))

@@

hc <- xx[,-1]  %>% 
  data.frame() %>% 
  `rownames<-`(xx$QSB) %>%  
  dist( ) %>% 
  hclust( method = 'ward')

xx %<>% mutate(comu.cl = as.numeric(cutree(hc,5)))

ddata_x <- dendro_data(as.dendrogram(hc))

ggplot(segment(ddata_x)) + 
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend),extent = "device") +
  geom_text(data = label(ddata_x), aes(label = label, x = x, y = -1.5)) + 
  coord_flip() + scale_y_reverse() + 
  theme(legend.position = 'none',axis.title.x = element_blank(),axis.title.y = element_blank()) + ylim(c(10,-3))

@
\caption{Dendrogam of communities based on importance measures. Communities close to each other have similar structure of key drivers of attachment. \label{dendro}}
\end{figure}

@@

\begin{table}[h]

>>>>

# to be removed...
# library(reshape2)
# library(plyr)

@@

<<data,echo=FALSE,warning=FALSE,message=FALSE, cache=FALSE>>=
# Procesing data sets from SOCT and Census Bureau
source('code/data_chk_tidy.R')
source('code/census_chk_tidy.R')

# loading results from cluster and random forest models.
res_cluster <- readRDS('data/res_cluster.rds') %>%
list2env(globalenv())

@@

# Compute comunity proportion of attached people
# community labels differ from soul and census datasets
lev <- soul %>% pull(QSB) %>% levels()

@@

rankings <- important %>%
  filter(rnk < 6) %>%
  group_by(dim, rnk) %>% 
  summarize(V1 = length(dim)) %>%
  ungroup() %>% 
  mutate(rnk = factor(rnk, labels = paste("R", 1:5, sep = "")))
@

@@

ggplot() + 
  geom_path(data = us, aes(long, lat, group=group), color = 'grey' ) +
  geom_point(data=lon.lat, aes(x=lon, y=lat, size = pop100), alpha = .5) +
  xlab('') + ylab('') + 
  scale_size(range = c(5,20)) + 
  scale_color_brewer(type = 'seq') +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(), 
        legend.position = 'none')
@
\caption{Geographical location of the 26 communities in the data, dot's size represents population. Only two communities on the west coast are included. \label{map}}
\end{figure}

@@

\begin{figure}[hbpt]
<<corplot, echo=FALSE, dependson=c('data'),cache=TRUE, warning = FALSE>>=
# With Census data we want to describe communities
cors <- census.data %>% 
  gather(vars, vv, -QSB, -CCEGRP2Attached, -`CCEGRP2Not Attached` ) %>%
  group_by(vars) %>% 
  summarise(
    cor.att = cor(vv, CCEGRP2Attached, use = 'complete.obs'),
    cor.noatt = cor(vv, `CCEGRP2Not Attached`, use = 'complete.obs') ) %>% 
  filter(!is.na(cor.att), vars != 'unemployed') %>% 
  filter(rank(cor.att) %in% c(1:10, 52:61) ) %>% 
  mutate(vars =  fct_reorder(vars, cor.att) , vars = recode_factor(
      vars, 
      "av.own" = 'Aver.years.own.in.community',
      "pop100"= 'size.community', 
      "av.rent"= 'Average.years.rent.in.community',
      "ow.prop10" = 'Prop.owner.more.10.years',
      "prop.female" = 'Pro.female',
      "prop2.1worker" = 'Prop.house.2.per.1.worker',
      "prop.4.per" = 'Prop.house.4.per.or.more',
      "ren.prop10" = 'Prop.rent.more.10.years',
      "own.rate" = 'Prop.own',
      "un.rate" = 'Unemployment.rate',
      "prop1.worker" = 'Prop.1worker',
      "Bachelor.s.degree.or.higher." = 'Prop.bachelor',
      "prop3.3worker" = 'Prop.house.3.per.3.workers',
      "prop.nat.entry.90s" = 'Prop.native.entry.90s',
      "prop.nat.entry.80s" = 'Prop.native.entry.80s',
      "militar" = 'Prop.militar',
      "prop1821" = 'Prop.pop.bw.18.21',
      "prop.entry.80s" = 'Prop.new.entry',
      "prop.male" = 'Prop.male',
      "prop.2.per" = 'Prop.2perhouse')) %>% 
  gather(type, vv, -vars) %>% 
  mutate( type = factor(type, labels = c('Attached', 'Not Attached') ),
          positive = factor(sign(vv), labels = c( 'Negative', 'Positive') ) )

cors %>%
  ggplot() +
  geom_point(aes(y = vars, x = vv, color = positive), size = I(3)) +
  facet_grid(~type) + theme(legend.position = 'bottom', legend.title = element_blank(), axis.text.y = element_text(size = I(15))) + labs(y = "", x = "Correlation")
@
\caption{Correlations with attached and not attached proportion in order from top to bottom based on correlation with attached\label{corplot}. This plot shows the 10 more relevant variables with positive correlation and 10 with negative correlation with attachment.}
\end{figure}

@@

cors %>%  
  filter(vars %in% c("Prop.new.entry", "Average.years.rent.in.community", "Aver.years.own.in.community") &  type =='Attached' ) %>% 
  select(vars, vv) %>% rename(cor = vv) %>%
  bind_rows(data.frame(vars = 'stay', cor = rho)) %>% 
  mutate(Data =  c(rep('Census Bureau', 3), 'Knight-Gallup data'), 
         Description = c( 'Average years owning the current home', 
                          'Average years renting in current home', 
                          'Prop. of people not borned in the community', 
                          'Proportion of age in current comunity')) %>% 
  rename(Correlation = cor) %>% 
  select(Data, Description, Correlation) %>% 
  xtable( caption='Correlation with attachment', label='lifetab') %>% 
  print( caption.placement = 'top', include.rownames = FALSE)
@

@@

\begin{figure}
<<life_seq, dependson=c('life'), echo=FALSE,fig.height=4,message=FALSE,warning=FALSE>>=
life %<>%
  mutate(QSB = factor(QSB, levels = rev(levels(QSB))))

@@

aux <- soul$QD111
levels(aux)[c(4,6:10,15:17,19,20) ] <- 'Other'
levels(aux)[c(1,2,5,7,8,10) ] <- 'Non Response'
levels(aux)[2] <- 'Hispanic'
aux[soul$QD10=='Yes'] <- 'Hispanic'
soul$race  <- factor(aux)

@@

qplot(data=tab, diver,logit(statistic.CCEGRP2Attached),size=I(5)) + theme(axis.text.y = element_text( size=I(10)) ) + ylab('Proportion of Attach (logit scale)') + xlab('Race diversity index') + geom_smooth(se=FALSE, method='lm', size=I(1))

@@

hc <- xx[,-1]  %>% 
  data.frame() %>% 
  `rownames<-`(xx$QSB) %>%  
  dist( ) %>% 
  hclust( method = 'ward.D')

# xx %<>% mutate(comu.cl = as.numeric(cutree(hc,5))) # what is this for ?
data.dendro <- hc %>% as.dendrogram() %>% dendro_data()

data.dendro %>% segment() %>% 
  ggplot() + 
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend),extent = "device") +
  geom_text(data = ggdendro::label(data.dendro), aes(label = label, x = x, y = -1.5)) + 
  coord_flip() + scale_y_reverse() + 
  theme(legend.position = 'none',axis.title.x = element_blank(),axis.title.y = element_blank()) + ylim(c(10,-3))
@
\caption{Dendrogam of communities based on importance measures. Communities close to each other have similar structure of key drivers of attachment. \label{dendro}}
\end{figure}

@@

\begin{table}[hbpt]
